---
title: "EDA"
author: "J L"
date: "October 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(viridis)
library(ggridges)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
```



## Load data
```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2016-01-01",
                      date_max = "2016-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, date, month, everything())
```
lubridate is tidyverse adjacent, but not loaded every time automatically when we're librarying tidyverse. So we're using "lubridate::floor date" here above


## group_by and counting
(Warning: group_by change behaviors of key functions, and may cause problems that's hard to debug.)
Because these (and other) functions will use grouping information if it exists, it is sometimes necessary to remove groups using ungroup().

Does group_by stay in the code chunk or global?
--only specific in code chunk, not globally.


```{r}
weather_df %>%
  group_by(name, month)
```


```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n = n())
```
n() counts the number of observations in the group


```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_unique = n_distinct(date)
    )
```
(Jeff doesn't use n_distinct a lot)


If we just need to count things, we don't need to group_by() & summarize(), the shortcut is to just use count()---But it's up to your habits
```{r}
weather_df %>%
  count(month, name = "n_days")
```


why table() in base R is not a good choice:
```{r}
weather_df %>%
  pull(name) %>% 
  table
```
This does not produce a dataframe that we can use afterwards


Let's make a nice table
```{r}
weather_df %>% 
  count(name) %>% 
  knitr::kable()
```



## (2*2 tables)

a digression ..

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE ~ ""              ## everything else, assign ""
    )
  ) %>% 
  group_by(name, cold) %>% 
  count() %>% 
  pivot_wider(
    names_from = cold,
    values_from = n
  )
```


Another way to do this using janitor::tabyl()
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE ~ ""              ## everything else, assign ""
    )
  ) %>%
  janitor::tabyl(name, cold)
```
(janitor has lots of useful functions)



## General summaries

```{r}
weather_df %>%
  group_by(name) %>%
  summarize(
    n = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),   
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```
na.rm = TRUE -- remove the missing values?
Do not set na.rm = TRUE by default! Check the missing values first before removing them.


```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax))
```


```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
    ) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```


Same table, but instead of making a plot, we're making this table more readable (if we're providing only this table and not the plot for others). As the original table contains 36 rows, not very readable.
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax), na.rm = TRUE) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```



## Grouped mutate

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(mean_tmax = mean(tmax, na.rm = TRUE))
```
This is mutating in a group-specific fashion


```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean(tmax)
    ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```



## Window functions in grouped mutates

```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax))
```


```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax)) %>% 
  filter(min_rank(tmax) == 1)   ##(or <2, same result)
```


If we want to show in order, from hottest to coldest day
```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(desc(tmax)) < 4)
```


* lags and leads (useful for using existing values to predict future ones)

```{r}
weather_df %>% 
  group_by(name) %>%
  mutate(
    lagged_tmax = lag(tmax)
  )
```

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    temp_change = tmax - lag(tmax)
    )
```

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(temp_change_sd = sd(temp_change, na.rm = TRUE),
            temp_change_max = max(temp_change, na.rm = TRUE))
```



## Limitations
summarize() can only be used with functions that return a single-number summary. This creates a ceiling, even if it is very high. Later we’ll see how to aggregate data in a more general way, and how to perform complex operations on the resulting sub-datasets.


