---
title: "Visualization_2"
author: "J L"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
library(patchwork)
```

## Create the weather data
```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```
cache: save the dataset to the directory, not redownloading it every time

## Initial plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5)
```

## Add labels
captions and titles
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minumum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  )
```

ticks and pokes (highlight scale on axis)
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 (too cold)", "0", "15")
    )
```

sqrt trans
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ºC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

## Colors

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(
    name = "Location", 
    h = c(100, 300)
    )
```

```{r}
ggp_base = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
    ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
    )
```

## Themes

```{r}
ggp_base +
  theme_bw() +
  theme(legend.position = "bottom")

ggp_base +
  theme_minimal() +                         ##order matters
  theme(legend.position = "bottom")   
```
theme_bw()--black and white theme
theme_minimal has to go first, otherwise it will not change the position to "bottom"

(Where do my warnings come from)
```{r}
weather_df %>% 
  filter(is.na(tmin))
```

* Setting options
In addition to figure sizing, I include a few other figure preferences in global options declared at the outset of each .Rmd file (this code chunk just gets copy-and-pasted to every new file).

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

There are ways to set color preferences globally as well (for example, to use viridis color palettes everywhere), although they’re a bit more involved.

## more than one dataset

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() +
  geom_line(data = central_park)
```

(breif aside about colors)
```{r}
waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = "red")) + geom_point()

waikiki %>% 
  ggplot(aes(x = date, y = tmax)) + geom_point(alpha = .5, color = "red")
```

## Multi-panel plots

```{r}
ggp_scatter = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_density = weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_density()

ggp_box = weather_df %>% 
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()
```

use patchwork package, to put plots together
```{r}
(ggp_scatter + ggp_density) / ggp_box
```

## Data manipulation

factor variables
```{r}
 weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_relevel(name, "Waikiki_HA", "CentralPark_NY")
  ) %>% 
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()

 weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()
```

## restructure then plot

when creating scatterplot, in order to reorder the plot (bringing "Central Park" to the front intead of "Waikiki"), we have to reorder the dataset instead of only using "fct_reorder" (factor_ordering)
```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = observation)) + 
  geom_density(alpha = .5) +
  facet_grid(~name) +
  theme(legend.position = "bottom")
```

As a final example, we’ll revisit the FAS data. We’ve seen code for data import and organization and for joining the litters and pups data. Here we add some data tidying steps to view pup-level outcomes (post-natal day on which ears “work”, on which the pup can walk, etc) across values of dose category and treatment day.
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, day_of_tx, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```
a more data-tidying problem rather than ggplot problem

