---
title: "Visualization_2"
author: "J L"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
library(patchwork)
```

## Create the weather data
```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```
cache: save the dataset to the directory, not redownloading it every time

## Initial plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5)
```

## Add labels
captions and titles
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minumum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  )
```

ticks and pokes (highlight scale on axis)
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 (too cold)", "0", "15")
    )
```

sqrt trans
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ºC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

## Colors

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(
    name = "Location", 
    h = c(100, 300)
    )
```

```{r}
ggp_base = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
    ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
    )
```

## Themes

```{r}
ggp_base +
  theme_bw() +
  theme(legend.position = "bottom")

ggp_base +
  theme_minimal() +                         ##order matters
  theme(legend.position = "bottom")   
```
theme_bw()--black and white theme
theme_minimal has to go first, otherwise it will not change the position to "bottom"

(Where do my warnings come from)
```{r}
weather_df %>% 
  filter(is.na(tmin))
```

## more than one dataset

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() +
  geom_line(data = central_park)
```

(breif aside about colors)
```{r}
waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = "red")) + geom_point()
```


