---
title: "Visualisation_1"
author: "J L"
date: "9/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```

## Lecture
  "gg" stands for "grammar of graphics"

## Create the weather data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Create a ggplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

alternate way of making this plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

adding color
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4)
```
  alpha--transparency

Why do `aes` positions matter?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4) +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```
  se = standard error?
  The former add (color = name) only to geom_points, the latter add (color = name) to the whole graph?

facet!
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE) +
  facet_grid(~name)
```
  --Good for individual reading, bad for comparison
  
Above is fine, but only limited to min & max values (should always be linear?)
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(size = 2, se = FALSE)

weather_df %>% 
  ggplot(aes(x = date, y = prcp, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

2d density
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_hex() +
  facet_grid(~name)
```
  geom_bin2d() similar to geom_hex()
  
## More kinds of plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, color = name)) +
  geom_histogram()
```
  This does not look right. For histogram, there is a difference between "color" and "fill"
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram()
```
  
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(~name)
```
  
density plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```

boxplots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

violin plots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

ridge plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

## Saving a plot
ex. ridge plot
```{r}
ggp_ridge_temp = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf", ggp_ridge_temp)
```

```{r, fig.width = 12}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```

